---
title: "Visualize Data"
output:
  html_document:
    df_print: paged
---

<!-- This file by Shannon Dunnigan was adapted from the orignal work at https://github.com/rstudio/master-the-tidyverse by RStudio. -->

```{r setup}
library(tidyverse)
library(here)
```


```{r loaddata}
wq <- read.csv(here::here('data', 'monthly_wq.csv')) 

```

## Quiz

What relationship do you expect to see between temperature (temp) and dissolved oxygen (do_mgl)?

## Your Turn 1

Run the code on the slide to make a graph. Pay strict attention to spelling, capitalization, and parentheses!

```{r}
ggplot(data = wq) +
  geom_point(mapping = aes(x = temp, y = do_mgl))

```

## Your Turn 2

Add `color`, `size`, `alpha`, and `shape` aesthetics to your graph. Experiment.  

```{r}
# original
ggplot(data = wq) + 
  geom_point(mapping = aes(x = temp, y = do_mgl))

# color
ggplot(data = wq) + 
  geom_point(mapping = aes(x = temp, y = do_mgl, color = state))

# size
# note that this is a continuous variable
ggplot(data = wq) + 
  geom_point(mapping = aes(x = temp, y = do_mgl, size = sal))

# alpha
# note that this is a continuous variable
ggplot(data = wq) + 
  geom_point(mapping = aes(x = temp, y = do_mgl, alpha = turb))
```

## Your Turn 3

Using the `wq` data, plot temperature (`temp`) by `month` as a scatterplot and facet by `state` and `coast`. 
Play around with `facet_grid()` and `facet_wrap()`. 

*What kinds of things do you notice?*

```{r}
ggplot(data = wq) +
  geom_point(mapping = aes(x = month, y = temp)) +
  facet_grid(. ~ state)

ggplot(data = wq) +
  geom_point(mapping = aes(x = month, y = temp)) +
  facet_wrap(. ~ state)

```

## Your Turn 4

Replace this scatterplot with one that draws boxplots. Use the cheatsheet. Try your best guess.

```{r}
ggplot(data = wq) +
  geom_point(mapping = aes(x = coast, y = temp)) 

# boxplots
ggplot(data = wq) +
  geom_boxplot(mapping = aes(x = coast, y = temp))

# geom_boxplot also has aesthetics you can map
ggplot(data = wq) +
  geom_boxplot(mapping = aes(x = coast, y = temp, color = state))

ggplot(data = wq) +
  geom_boxplot(mapping = aes(x = coast, y = temp, fill = state))

ggplot(data = wq) +
  geom_boxplot(mapping = aes(x = coast, y = temp, linetype = coast))

```

## Your Turn 5

Load the `diamonds` dataset and examine it. *ahem* Remember the `glimpse()` function? This data set contains prices and other attributes of diamonds and is provided by the `ggplot2` package.

Create a bar chart of cut colored by cut. Use the cheatsheet. 

```{r}
diamonds

# always good to check your data
glimpse(diamonds)

ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, color = cut)) # color does not "fill" the bars

ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, fill = cut)) # must use fill to shade in the full bar


```

## Your Turn 6

Determine what the default stat is for each of the following geoms:

`geom_line`
`geom_density`
`geom_smooth`

```{r}
# be sure to check the help pane
?geom_line
?geom_density
?geom_smooth

```

## Your Turn 7

Predict what this code will do. Then run it.

```{r}
ggplot(wq) +
  geom_point(aes(temp, do_mgl)) +
  geom_smooth(aes(temp, do_mgl)) 

```

## Your Turn 10 

What does `getwd()` return?

```{r}
getwd() # returns your working directory (i.e., where R is looking for and saving your files.)
```


## Your Turn 11

Save the last plot and then locate it in the files pane.

```{r}
ggsave("my-last-plot.png", dpi = 130)
```

## MINI CHALLENGE

Using the diamonds data make a scatterplot of carat by price and color by clarity. 
Customize the themes and colors of the plot to make it the *ugliest* plot that you can. 

``` {r ugly}
ggplot(data = diamonds) +
  geom_point(aes(x = carat, y = price, color = clarity))
```

***

# Take aways

You can use this code template to make thousands of graphs with **ggplot2**.

```{r, eval = FALSE}
ggplot(data = <DATA>) +
  <GEOM_FUNCTION>(mapping = aes(<MAPPINGS>),
                  stat = <STAT>,
                  position = <POSITION>) +
  <COORDINATE_FUNCTION> +
  <FACET_FUNCTION> +
  <SCALE_FUNCTION> +
  <THEME_FUNCTION>
```
