---
title: "Visualize Data"
output:
  html_document:
    df_print: paged
---

<!-- This file by Shannon Dunnigan was adapted from the orignal work at https://github.com/rstudio/master-the-tidyverse by RStudio. -->

```{r setup}
library(tidyverse)
library(here)
```


```{r loaddata}
wq <- read.csv(here::here('data', 'monthly_wq.csv')) 

dplyr::glimpse(wq)
```

## Quiz

What relationship do you expect to see between temperature (temp) and dissolved oxygen (do_mgl)?

## Your Turn 1

Run the code on the slide to make a graph. Pay strict attention to spelling, capitalization, and parentheses!

```{r}


```

## Your Turn 2

Add `color`, `size`, `alpha`, and `shape` aesthetics to your graph. Experiment.  

```{r}
ggplot(data = wq) + 
  geom_point(mapping = aes(x = temp, y = do_mgl, 
                           color = state, shape = coast,
                           alpha = temp),
             size = 5)

```

## Your Turn 3

Using the `wq` data, plot temperature (`temp`) by `month` as a scatterplot and facet by `state` and `coast`. 
Play around with `facet_grid()` and `facet_wrap()`. 

*What kinds of things do you notice?*

```{r}
ggplot(data = wq) +
  geom_point(mapping = aes(x = month, y = temp, color = coast)) +
  facet_wrap( ~ station_code)

```

## Your Turn 4

Replace this scatterplot with one that draws boxplots. Use the cheatsheet. Try your best guess.

```{r}
ggplot(data = wq) +
  geom_point(mapping = aes(x = coast, y = temp)) 

```

## Your Turn 5

Load the `diamonds` dataset and examine it. *ahem* Remember the `glimpse()` function? This data set contains prices and other attributes of diamonds and is provided by the `ggplot2` package.

Create a bar chart of `cut` colored by `cut`. Use the cheatsheet. 

```{r}
diamonds

ggplot(data = diamonds)+
  geom_bar(mapping = aes(x = cut, color = cut))

ggplot(data = diamonds)+
  geom_bar(mapping = aes(x = cut, fill = cut))

```

## Your Turn 6

Determine what the default stat is for each of the following geoms:

`geom_line()`
`geom_density()`
`geom_smooth()`

```{r}
?geom_point

```

## Your Turn 7

Predict what this code will do. Then run it.

```{r}
a <-ggplot(wq) +
  geom_point(aes(temp, do_mgl)) +
  geom_smooth(aes(temp, do_mgl)) 
a
```

## Your Turn 8 

What does `getwd()` return?

```{r}
getwd()
```

## Your Turn 9

Save the last plot and then locate it in the files pane.

```{r}
ggsave("do_by_temp.png", width = 6, height = 4, units = "in")
```

## MINI CHALLENGE

Using the `diamonds` data make a scatterplot of `carat` by `price` and assign color by `clarity`. 
Customize the themes and colors of the plot to make it the *ugliest* plot that you can. 
Save this plot in your files pane.
```{r ugly}
ggplot(data = diamonds) +
  geom_point(mapping = aes(x = carat, y = price, colour = clarity)) +
  scale_color_brewer(type = "div", palette = "RdGy") +
  theme(panel.background = element_rect(color = "purple"),
        axis.text = element_text(size = 20), 
        plot.background = element_blank())
```


***

# Take aways

You can use this code template to make thousands of graphs with **ggplot2**.

```{r, eval = FALSE}
# ggplot(data = <DATA>) +
#   <GEOM_FUNCTION>(mapping = aes(<MAPPINGS>),
#                   stat = <STAT>,
#                   position = <POSITION>) +
#   <COORDINATE_FUNCTION> +
#   <FACET_FUNCTION> +
#   <SCALE_FUNCTION> +
#   <THEME_FUNCTION>
```
