---
title: "Final Challenge"
author: "Kim Cressman"
output: 
    html_document:
        toc: true
        toc_depth: 3
        toc_float:
            collapsed: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Setup:

```{r, message = FALSE, warning = FALSE}
library(dplyr)
library(tidyr)
library(lubridate)
library(here)
library(ggplot2)
```



# Final challenge  

For our final challenge, you are going to put together the things that you've learned throughout the day. You will write a loop to read in each of a few files, and produce a graph and two summary tables (one needs to be pivoted) for each.  


## 1. Build up vector of file names

We can list the files in a directory using the `dir()` function:     

```{r}
dir(here::here("data", "final_challenge"))
```

If you only want files that include a certain string, you can include a `pattern` argument (for example, if you had mutiple file types in the directory but only wanted to include CSVs; or if you only wanted to include certain stations; etc.)  

```{r}
dir(here::here("data", "final_challenge"), pattern = ".csv")
```


Assign those to a vector named `my_files`. This is the file vector that we'll loop through.  

```{r}

```



## 2. Write functions  

Write a function to calculate standard error. Name it `sterr`. Recall that the body of the function should look something like this:  

```{r}
sd( data, na.rm = TRUE) / sqrt( sum(!is.na( data )) )

sterr <- ---------
```

Write another function to calculate depth in feet, from meters. (Yes, I know this math is probably easier to just *do* in a call to `mutate`. Humor me and practice your function-writing.) 

The conversion is:  
`feet = meters * 3.28`  

```{r}
m_to_ft <- ---------
```



## 3. Fill in the loop structure  


```{r}
for(i in -------(-----)){
    
    # find the file
    in_file <- here::here("data", "final_challenge", --------)
    
    # read in the file
    dat <- ------(in_file, stringsAsFactors = FALSE)
    
    
    
    # wrangle the data frame a bit
    dat <- dat %>% 
        mutate(
            # fill in your function to turn depth in m into depth in ft
            depth_ft = -----(----),
            # fill in the function needed to make sure date is correctly "seen"
            date = lubridate::----(------),
            # the lines below are "gimmes" because we haven't discussed them
            # but: they pull out the portion of the date we want, and 
            # we can use that later
            year = lubridate::year(date),
            month = lubridate::month(date))
    
    
    
    # make AND PRINT a line graph of depth in feet on the y-axis, 
    # by date on the x-axis, and where each year is a different color 
    # add in labels and themes if you want to make it look nice!
    
    p <- ggplot(dat) +
        ---------
    
    
        # did you remember to print the result?
        
        
    
    # now generate and print a summary table, grouped by year,
    # to show median, mean, and sterr of depth_ft
    
    # hint: if you use a filter statement to get rid of 
    # rows where depth_ft is NA  (in the dplyr script we learned how
    # to negate "is.na" ), you don't have to worry about
    # using "na.rm = TRUE" in each individual summary function
    dat %>% 
        filter(!is.na(depth_ft)) %>%
        -----------
        
        
        
        # did you remember to print the result?
        
        
        
    
    # Group by year and month, calculate mean of depth_ft,
    # and pivot the data frame so you get
    # a table with one row per year, and one column per month
    # the values in each cell are the mean depth for that month/year combination
    
    
    dat %>% 
        --------------
        
        
        # did you remember to print the result?
        
}
```


